---
title: "R Notebook"
output: html_notebook
---

2. Recall the marriage, age, and happiness collider bias example from Chapter 6. Run models m6.9 and m6.10 again. Compare these two models using
WAIC (or LOO, they will produce identical results). Which model is expected to make better predictions? Which model provides the correct causal
inference about the influence of age on happiness? Can you explain why the
answers to these two questions disagree?

H -> M <- A

```{r}
library(rethinking)
d <- sim_happiness( seed=1977 , N_years=1000 )
precis(d)
```


Consider only adults and rescale age
```{r}
d2 <- d[ d$age>17 , ] # only adults 6.23
d2$A <- ( d2$age - 18 ) / ( 65 - 18 )
```


COnsider the following model to infer the influence of age on happiness (note: MArriage is a collider so it'll trick us)
```{r}
d2$mid <- d2$married + 1 # marriage status index
m6.9 <- quap(
alist(
happiness ~ dnorm( mu , sigma ), # target
mu <- a[mid] + bA*A,
a[mid] ~ dnorm( 0 , 1 ), # PRIOR: When A = 0, i.e. age = 18 then happiness mu = a. so a needs to cover the hapiness range (i.e. -2:2), so std(a) = 1
bA ~ dnorm( 0 , 2 ), # PRIOR: Std = 2 ensures that the whole happiness range (i.e. -2:2 = 4) span by the strongest bA*A relationship (i.e. bA = 1) is withing 2sd e.g. 95% of mass
sigma ~ dexp(1)
) , data=d2 )
precis(m6.9,depth=2)
```

The model believes that happiness in negatively associated with happiness.

However, now we compare it with this other model omitting marriage status:
```{r}
m6.10 <- quap(
alist(
happiness ~ dnorm( mu , sigma ),
mu <- a + bA*A,
a ~ dnorm( 0 , 1 ),
bA ~ dnorm( 0 , 2 ),
sigma ~ dexp(1)
) , data=d2 )
precis(m6.10)
```

This model finds no association between happiness and age. This is a typical collider situation.
Age association with happiness is spuios not causal.


```{r}
set.seed(77)
compare(m6.9, m6.10, func = LOO)
compare(m6.9, m6.10, func = WAIC)
```

WAIC: approximation of lppd deviance
m6.9 is the best model FOR PREDICION. THe spurious correlation adds information and helps prediction accuracy. However, if we were interested in causality (e.g. because we want to intervene in the behaviour) we'd have to chose 6.10

pWAIC: penalty term
close to number of parameters

dWAIC:
is the difference between each model's WAIC and the best WAIC in the set.

SE and dSE
standard error of WAIC and dWAIC